---
# Choose from this list of Journals:
# JGR: Atmospheres, JGR: Biogeosciences, JGR: Earth Surface, 
# JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Space Physics, 
# Global Biogeochemical Cycles, Geophysical Research Letters, 
# Paleoceanography and Paleoclimatology, Radio Science, Reviews of Geophysics, 
# Tectonics, Space Weather, Water Resources Research, Geochemistry, Geophysics, 
# Geosystems, Journal of Advances in Modeling Earth Systems (JAMES),
# Earth's Future, Earth and Space Science, Geohealth
journal: "Water Resources Research"
# Use draft to submit a paper
classoption: "draft,linenumbers"
# A title should be specific, informative, and brief. Use
# abbreviations only if they are defined in the abstract. Titles that
# start with general keywords then specific terms are optimized in
# searches
title: "True Water Level Uncertainty"
# First name or initial followed by last name
# Authors are individuals who have significantly contributed to the
# research and preparation of the article. Group authors are allowed, if
# each author in the group is separately identified in an appendix.
# Additional author notes should be indicated with
authors:
  - name: Joseph P. Shannon
    affil: 1
    thanks: Joe's Thanks
  # - name: Fred T. Secondauthor
  #   affil: 1
  #   thanks: "Current address: Some other place, Germany"
  # - name: I. Ken Groupleader
  #   affil: "1, 2"
affiliations:
  - number: 1
    name: "Michigan Technological University, College of Forest Resources and Environmental Science"
  # - number: 2
  #   name: "The second affiliation"
# More than one corresponding author is allowed in this LaTeX file and for
# publication; but only one corresponding author is allowed in our editorial system.
corresponding_author:
  - name: Joseph P. Shannon
    email: jpshanno@mtu.edu
keypoints:  
  - "List up to three key points (at least one is required)"
  - "Key Points summarize the main points and conclusions of the article"
  - "Each must be 100 characters or less with no special characters or punctuation"
abstract: "A good abstract will begin with a short description of the problem being addressed, briefly describe the new data or analyses, then briefly states the main conclusion(s) and how they are supported and uncertainties."
plain_language_summary: "Some journals require a plain language summary. See: https://publications.agu.org/author-resource-center/text-requirements/#abstract"
output: 
    rticles::agu_article
bibliography: "`r rbbt::bbt_write_bib('biblio.bib', translator = 'bibtex', overwrite = TRUE)`"
header-includes: 
      - \usepackage{soulutf8}  # For UTF8 chars in TrackChanges
# AGU recommends using the trackchanges LaTeX package in the edition process
# which is available from this link:
# https://publications.agu.org/files/2019/02/January-14-2019-latex-templates.zip
---

```{r echo=FALSE}
# Some recommended settings. 
knitr::opts_chunk$set(
  echo = FALSE,
  fig.pos = 'h',
  out.extra = "",   # To force the use of figure enviroment
  fig.cap = "Please caption every figure"
)
```

```{r keypoints_check, echo=FALSE, results='asis', eval = TRUE}
# This chunk adds a warning if any keypoint is longer than 100 characters. 
# To disable it, you can remove it or set eval to FALSE.
if (any(nchar(rmarkdown::metadata$keypoints) > 100)) {
  cat("\\textcolor{red}{\\textbf{Warning}: keypoint(s)", 
      knitr::combine_words(which(nchar(rmarkdown::metadata$keypoints) > 100)), 
      "longer than 100 characters.}")
}
```

Suggested section heads

# Introduction

Much research in and monitoring of water levels for wells, wetland surface water,
and stream discharge __(sewer discharge and other applications too)__ utilize 
pressure-transducer measurements. There are known artifacts to these measurements
when the temperature of levelloggers is not addressed [@moore-2016] or when the
barometric logger and water logger are not deployed in similar thermal regimes
[@cuevas-2010; @mclaughlin-2011]. @moore-2016 suggests that all loggers be 
tested for inaccuracies over the range of temperature measurements. 
@mclaughlin-2011 found that .... These results are similar to @cuevas-2010, who
found a __%ERROR__ was possible when barometric and water loggers were deployed
in different thermal settings (__DOES THIS HOLD FOR EXTERNAL BAROMETRIC DATA?__).

@tamari-2010 highlighted important differences in how logger uncertainty can be
expressed

# Materials and Methods

We have conducted four experiments to determine the direct impact of temperature 
regimes on levellogger water and barometric measurements as well as temperature
differentials. The experiments represent a fully crossed design with stable and
variable temperatures and similar and dissimilar air and water thermal 
regimes. Eightteen levelloggers of varying age and model, all manufactured by
Solinst, Ltd. (Georgetown, ON, Canada), were used to recorded water levels (
__TABLE 1__). Two levelloggers, both Solinst Levellogger Juniors, were used to 
monitor barometric pressure. Water levels were recorded under a fixed water
level for compenstation, and barometric pressure was compared to nearby weather
stations.

Water levelloggers were suspended from a brace across a large container. The
levelloggers were suspended in a manner to align the zero point of each logger.
The loggers were covered with a fix depth of water and a layer of canola oil was
added to prevent evaporation over the length of the experiment. Water level was
measured at the start and end of experimental period. To provide a thermally
similar air and water environments we placed a graduated cylinder in the center
of the water. The cylinder was partially filled with gravel and water to
counteract bouyancy of they cylinder. During the two experiments with sychronous
thermal regimes the two barometric loggers were suspended directly above the
gravel surface. The entire set-up was covered loosely with aluminum foil to
prevent direct wind effects. External barometric pressure was retreived for two
nearby locations accessed via the Synoptic Mesonet API
(https://developers.synopticdata.com/mesonet/v2/). For stable thermal regimes
the experiments were conducted in a non-climate controled  basement to minimze
temperature fluctions. Varying thermal regime experiments were conducted
outdoors out of direct solar radiation.

In order to test the compensation models developed using the nested experiments 
the data collection was repeated under the varying, dissimilar conditions.

Levelloggers were deployed and data retreived using using Solinst Levellogger 
v. 4.4.0 software. Loggers were set to record at 1-minute intervals with 
time synchronized to computer time at launch. The four experiments were run 
serially without retreiving data from the loggers. A settling time of at least
30 minutes was observed after barometric loggers or entire experimental apparatus
were relocated.

<!-- Please use ONLY \\citet and \\citep for reference citations. -->
<!-- DO NOT use other cite commands (e.g., \\cite, \\citeyear, \\nocite, \\citealp, etc.). -->
<!-- Example \\citet and \\citep: -->
<!--   ...as shown by \citet{Levitus2012}, \citet{Nuncio2011} and \citet{Raphael2004} -->
<!--   ...as shown by \citep{Levitus2012}, \citep{Nuncio2011}, \citep{Raphael2004}. -->
<!--   ...has been shown \citep[e.g.,][]{Levitus2012, Nuncio2011, Raphael2004}. -->

# Data


# Results

There was a storm that came through, could have strong pressure gradient
One barologger found resting on gravel/in water in cylinder during variable,
similar experiment.

Or section title might be a descriptive heading about the results

Enter Figures and Tables near as possible to where they are first mentioned:
DO NOT USE \\psfrag or \\subfigure commands.
DO NOT USE \\newcommand, \\renewcommand, or \\def, etc.

```{r}
plot(pressure)
```

Example table

\begin{table}
 \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
 \centering
 \begin{tabular}{l c}
 \hline
  Run  & Time (min)  \\
 \hline
   $l1$  & 260   \\
   $l2$  & 300   \\
   $l3$  & 340   \\
   $h1$  & 270   \\
   $h2$  & 250   \\
   $h3$  & 380   \\
   $r1$  & 370   \\
   $r2$  & 390   \\
 \hline
 \multicolumn{2}{l}{$^{a}$Footnote text here.}
 \end{tabular}
 \end{table}

AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.

```{r, fig.env = "sidewaysfigure"}
plot(pressure)
```

\begin{sidewaystable}
\caption{Caption here}
\label{tab:signif_gap_clos}
\begin{tabular}{ccc}
one&two&three\\
four&five&six
\end{tabular}
\end{sidewaystable}

If using numbered lines, please surround equations with \\begin{linenomath\*}... \\end{linenomath\*}

\begin{linenomath*}
\begin{equation}
y|{f} \sim g(m, \sigma)
\end{equation}
\end{linenomath*}

# Conclusions

# References



\appendix
\section{Here is a sample appendix}

Optional Appendix goes here

Optional Glossary, Notation or Acronym section goes here:

Glossary is only allowed in Reviews of Geophysics

\begin{glossary}
\term{Term}
 Term Definition here
\term{Term}
 Term Definition here
\term{Term}
 Term Definition here
\end{glossary}


\begin{acronyms}
\acro{Acronym}
 Definition here
\acro{EMOS}
 Ensemble model output statistics
\acro{ECMWF}
 Centre for Medium-Range Weather Forecasts
\end{acronyms}

\begin{notation}
\notation{$a+b$} Notation Definition here
\notation{$e=mc^2$}
Equation in German-born physicist Albert Einstein's theory of special
relativity that showed that the increased relativistic mass ($m$) of a
body comes from the energy of motion of the body—that is, its kinetic
energy ($E$)—divided by the speed of light squared ($c^2$).
\end{notation}


\acknowledgments
The acknowledgments must list:
A statement that indicates to the reader where the data
supporting the conclusions can be obtained (for example, in the
references, tables, supporting information, and other databases).

All funding sources related to this work from all authors

Any real or perceived financial conflicts of interests for any author

Other affiliations for any author that may be perceived as having a conflict of 
interest with respect to the results of this paper.

It is also the appropriate place to thank colleagues and other contributors.

AGU does not normally allow dedications. 
