---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Magnitude, Consequences, and Correction of Temperature-Derived Errors for Absolute Pressure Transducers under Common Monitoring Scenarios

This repository contains the code and data for the above-titled manuscript 
submitted for publication to Hydrological Processes, September, 2021.

### Abstract
Continuous water level monitoring using absolute pressure transducers with
onboard datalogging is common practice in hydrologic studies. While there has
been some discussion and study of temperature-derived error (TDE), there has not
been a systematic evaluation of the problem. We sought to answer three
questions: 1) are current best-practices enough to avoid these errors, 2) can
laboratory correction be used to correct field data from varying conditions, and
3) what is the scale of the additional uncertainty of the correction procedure?
We evaluated the magnitude of such errors under laboratory conditions that
mimicked common monitoring scenarios. Using field data, we also demonstrated the
impact of TDEs on calculated daily mean water level and diurnal signal
decomposition to estimate evapotranspiration (ET). To address instrument and
model uncertainty, we fit 1000 possible correction models using a
double-bootstrap approach. Correction models fit expected error as a function of
water and air temperature and rate of change of air temperature. TDEs were a
significant source of error, resulting in recorded data outside of
manufacturer-stated instrument uncertainty, with 45% of bootstrap models showing
significant but small TDEs under best-practice deployment. Correction equations
did introduce additional error, often on a much smaller scale than instrument
uncertainty. When tested against a validation dataset, correction equations
effectively reduced total measurement uncertainty below instrument uncertainty
by up to 65%. The effects of TDEs on case-study field data resulted in 56% of
daily mean values outside of instrument error bounds (errors: -1.5--4.2 cm).
Our results suggest that a single laboratory correction equation can be used
across monitoring scenarios, though we suggest matching deployment conditions as
closely as practical. Identification and correction of TDEs are essential to
avoid misguided conclusions, downstream analysis, and water resources
management.

### Keywords
Pressure transducer, water level correction, measurement uncertainty,
groundwater monitoring, streamflow measurement, stage

### Graphical Abstract

```{r, echo = FALSE}

knitr::include_graphics("output/manuscript_figures/Figure_6-Case_Study_Panel.tiff")

```